Automated Verification of Stand-alone Solar Photovoltaic Systems: Optimal Sizing and Project Validation

### 2019-10-29 Tex/Old text/abstract.tex
Thesis update

\thispagestyle{plain}
\begin{center}
%    \Large
%    \textbf{Automated Verification Applied to Stand-alone Solar Photovoltaic Systems}
%    
%    \vspace{0.4cm}
%    \large
%    Optimal Sizing and Project Validation
%    
%    \vspace{0.4cm}
%    \textbf{Alessandro Trindade}
%    
    \vspace{0.9cm}
    \textbf{Abstract}
\end{center}
With declining costs and increasing performance, the deployment of renewable energy systems is now growing faster than in the previous decade. In 2017, for the first time, the number of people without access to electricity dipped below 1 billion, but trends on energy access likewise fall short of global goals. Particular attention is given to stand-alone solar photovoltaic systems in rural areas or where grid extension is unfeasible. Tools to evaluate or to size electrification projects are available, but they are based on simulations that do not cover all aspects of the design space. However, the use of formal methods to model and validate any system is growing with time, mainly to find bugs in sophisticated hardware and software systems: they aim to establish system correctness with mathematical rigor. The use of formal methods in electrical systems is a new subject, with research being published only in the last four years.
Moreover, the use of automated synthesis in order to obtain the optimal sizing of solar photovoltaic systems was never done before. This PhD thesis marks two mains achievements. First, the application of software model checking to formally verify the design of a stand-alone solar photovoltaic system, including solar panel, charge controller, battery, inverter, and electric load. Second, a sound and automated approach to obtain optimal sizing of stand-alone photovoltaic systems using program synthesis. Here every component and function of a solar photovoltaic system is described, including its properties and the behavioral model that represents the optimal sizing. Related to formal verification, case studies, from real photovoltaic systems deployed in five different sites, ranging from 700 W to 1,200 W, are used to evaluate the proposed approach and to compare that with specialized simulation tool. Different verification tools are evaluated as well in order to compare performance and soundness. Data from practical applications show the effectiveness of our proposed approach, where specific conditions that lead to failures in a photovoltaic solar system are only detailed by the automated verification method.
Moreover, concerning the use of program synthesis, we propose a variant of counterexample guided inductive synthesis (CEGIS) approach. This variant has two phases linking the technical and cost analysis. First, we synthesize a feasible candidate based on power reliability, but that may not achieve the lowest cost. Second, the candidate is then verified iteratively with a lower bound cost via symbolic model checking. If the verification step succeeds, the lower bound is adjusted; if it fails, a counterexample provides the optimal solution. The proposed synthesis method is novel and unprecedented to streamline the design of photovoltaic systems. Experimental results using seven case studies and financial equipment data show that our synthesis method can produce within an acceptable run-time the optimal system sizing. We also present a comparative with a specialized simulation tool over real photovoltaic systems to show the effectiveness of our approach, which can provide a more detailed and accurate solution than that simulation tool.

>>>>

\thispagestyle{plain}
\begin{center}
%    \Large
%    \textbf{Automated Verification Applied to Stand-alone Solar Photovoltaic Systems}
%    
%    \vspace{0.4cm}
%    \large
%    Optimal Sizing and Project Validation
%    
%    \vspace{0.4cm}
%    \textbf{Alessandro Trindade}
%    
    \vspace{0.9cm}
    \textbf{Abstract}
\end{center}
With declining costs and increasing performance, the deployment of renewable energy systems is now growing faster than in the previous decade. In 2017, for the first time, the number of people without access to electricity dipped below 1 billion, but trends on energy access likewise fall short of global goals. Particular attention is given to stand-alone solar photovoltaic systems in rural areas or where grid extension is unfeasible. Tools to evaluate or to size electrification projects are available, but they are based on simulations that do not cover all aspects of the design space. However, the use of formal methods to model and validate any system is growing with time, mainly to find bugs in sophisticated hardware and software systems: they aim to establish system correctness with mathematical rigor. The use of formal methods in electrical systems is a new subject, with research being published only in the last four years.
Moreover, the use of automated synthesis in order to obtain the optimal sizing of solar photovoltaic systems was never done before. This Ph.D. thesis marks two mains achievements. First, the application of software model checking to formally verify the design of a stand-alone solar photovoltaic system, including solar panel, charge controller, battery, inverter, and electric load. Second, a sound and automated approach to obtain optimal sizing of stand-alone photovoltaic systems using program synthesis. Related to formal verification, we used case studies from real photovoltaic systems deployed in five different sites, ranging from 700 W to 1,200 W, in order to evaluate the proposed approach and to compare that with specialized simulation tool. Different verification tools are evaluated as well in order to compare performance and soundness. Data from practical applications show the effectiveness of our proposed approach, where specific conditions that lead to failures in a photovoltaic solar system are only detailed by the automated verification method.
Moreover, concerning the use of program synthesis, we propose a variant of counterexample guided inductive synthesis (CEGIS) approach. This variant has two phases linking the technical and cost analysis. First, we synthesize a feasible candidate based on power reliability, but that may not achieve the lowest cost. Second, the candidate is then verified iteratively with a lower bound cost via symbolic model checking. If the verification step succeeds, the lower bound is adjusted; if it fails, a counterexample provides the optimal solution. The proposed synthesis method is novel and unprecedented to streamline the design of photovoltaic systems. Experimental results using seven case studies show that our synthesis method can produce within an acceptable run-time the optimal system sizing. We also present a comparative with a specialized simulation tool over real photovoltaic systems to show the effectiveness of our approach, which can provide a more detailed and accurate solution than that simulation tool.

### 2019-10-29 Tex/Old text/chapters/appendixB.tex
Thesis update

\textcolor{red}{the second part of this sentence is unclear...} In general, we pursuit to write code that could be verification tool independent, i.e., that can be used in any model checking tool without adaptation, since we aimed to evaluate different.

We wrote the tools in the C programming language, without the use of '\# include', or '\# define'. Based on the fact that there exists no use of compiler library, it was necessary to write some functions, which perform specific calculations according to the adopted mathematical models (e.g., logarithm of natural order and exponential).

@@

The only block that demands user intervention to input data is block 1 of both tools \textcolor{red}{where can I see block 1? which tools?}. If it is aimed to change the mathematical model adopted, so coding must be done in block 3 of the automated verification tool and block 2 of the automated synthesis tool.

>>>>

One premise of our tools was to be written as an ANSI C program, and to follow the requirements established for the International Competition on Software Verification (SV-COMP) \footnote{https://sv-comp.sosy-lab.org/}. That allow us to write a code that could be executed into three different verifiers without adaptation, and even in the future other verifiers can be used as well.
%
%\textcolor{red}{the second part of this sentence is unclear...} In general, we pursuit to write code that could be verification tool independent, i.e., that can be used in any model checking tool without adaptation, since we aimed to evaluate different.

Therefore, we wrote the code without the use of '\# include', or '\# define'. Moreover, based on the fact that there exists no use of compiler library, it was necessary to write some functions, which perform specific calculations according to the adopted mathematical models (e.g., logarithm of natural order and exponential).

@@

The only block that demands user intervention to input data is the block 1 listed in~\ref{sec:automatedverification} and \ref{sec:automatedsynthesis} from the automated verification and synthesis tools respectively. If it is aimed to change the mathematical model adopted, so coding must be done in block 3 of the automated verification tool and block 2 of the automated synthesis tool.

### 2019-10-29 Tex/Old text/chapters/introduction.tex
Thesis update

%%automated verification
\textcolor{red}{What is the relation between CPS and PV? We should make it clear already in the beginning...}
Cyber-Physical Systems (CPS) are engineered systems, which are built from, and depend upon, the seamless integration of computational and physical components~\cite{NSF2015,ChavesIBCF19}. During operation, such components must frequently adapt to the operating environment changes (dynamics of the physical processes) faced at run-time. They must be able to continue to behave in a controlled and safe way, thus posing novel technical challenges for the software engineering of services and applications for CPS~\cite{Metzger2014}. Software pervasiveness in CPS places new challenges; in particular, highly dynamic environments, rapidly changing requirements, unpredictable and uncertain operating conditions demand new paradigms for software design~\cite{Filieri2015}.

While some research efforts do exist to enhance and optimize the software development processes for CPS, further investigation and discussion of better and more effective models are still needed in practice~\cite{Al-Jaroodi2016}. Among the opportunities for enhancements in the software development processes for CPS, there exists the need for developing new techniques and tools to support CPS requirements gathering and analysis to synthesize \textit{correct-by-construction} implementations of CPS. These techniques need to deal with predefined requirements enforced by nature and inherited constraints of the target CPS. Besides, they should be able to provide verification and validation mechanisms for completeness, correctness, and consistency~\cite{Al-Jaroodi2016}. Uncertainty and variability, at the same time, can be dealt with formal verification~\cite{NESSI}. Energy production, distribution, and optimization are all CPS problems~\cite{UC}.

Lack of access to clean and affordable energy is considered a core dimension of poverty~\cite{Hussein2012}. Nevertheless, the world is progressing; in particular, the number of people without electricity access fell below 1 billion threshold for the first time in 2017~\cite{IEAweo2018}. The share of people without access to electricity from Africa is 58\%, while 19\% of the share comes from developing Asia, and 31\% from Latin America~\cite{IEAweo2018}. Numbers from Brazil show the aim to electrify 270 isolated places and 2.7 million people until 2023~\cite{EPE2018}. 
Furthermore, there exists a close relationship between the lack of energy and the low HDI (Human Development Index) of those localities~\cite{Coelho}. Therefore, increased access to energy allows economic growth and poverty alleviation \cite{Karekesi}.

@@

The main objective of this thesis is thus to propose and evaluate the use of formal methods in solar PV systems, using a computer science method to solve specific issues of the electrical engineering area. \textcolor{red}{This sentence is too long and unclear...} In particular, we focus on automated verification by model checking, on stand-alone solar PV systems that is most adequate to rural areas of developing countries, on the need to validate and optimize those systems, on evaluating different state-of-art symbolic software verifiers (and solvers) with the goal of obtaining the best performance in our verification back-end, on applications written in ANSI-C that are platform-independent; and we do not perform code modification on the verifiers or solvers, we use it as available by their developers. We further evaluate our proposed solution by comparative with commercial simulation tool and by data collected from real PV systems deployed at riverside communities.
%More specifically we will:
%The thesis produces a methodological research with innovative value regarding the first use of automated synthesis for optimal sizing of solar PV systems.

\textcolor{red}{A PhD thesis is not simply the application of methods and development of tools. Please rephrase this sentence.}
This Ph.D. thesis has the novelty of applying automated verification method to solar PV systems. The result is the development of two software: one to perform automated verification and validate the solar PV systems behavior, and another one to produce the optimal sizing of solar PV systems through automated synthesis program.

In this study, a mathematical model of each component of a stand-alone PV system, as panel solar, charge controller, batteries, inverter, and electrical load is used for formal verification. The behavior of each system component can be analyzed and observed with the support of those formal models, as a joint operation of the components, which in this case represents the operation of the solar PV system itself. The project requirements, as battery autonomy and power demand, besides weather conditions, as solar irradiance and ambient temperature, are provided to the proposed tool and automatically verified during the formal process. The model checking tool reports in which conditions a system does not meet the user requirements. A key benefit to this approach is that it helps in the detection of flaws in the design phase of system development, thereby considerably improving system reliability~\cite{Akram2018}. 
%Those programs are viewed as mathematical objects, and the use of 
%Bringing this technique to solar PV systems, i
%Instead of finding a bug or a software inconsistence, t
%Here it will be verified a stand-alone PV system, with all their components modeled mathematically,   
%Among the non-deterministic variables considerated, is it worth to mention environmental issues (like solar irradiance, and temperature), and the variation of the electrical load (power consumption) or the battery' state of charge, as shown in Fig.\ref{fig:flowchartgeneral}.
\textcolor{red}{This paragraph is too long and unclear...}
The implementation of the proposed tool is carried out by means of an algorithm implemented using the C programming language, which is executed by three state-of-the-art model checkers to formally verifying PV designs, in order to evaluate performance and correctness: the C Bounded Model Checker (CBMC)~\cite{Kroening}, the Efficient SMT-based Bounded Model Checker (ESBMC)~\cite{esbmc2018}, and the Configurable Program Analysis Checker (CPAchecker).
%three the efficient SMT-based bounded model checker (ESBMC)~\cite{esbmc2018}. %, which allows one to incrementally verify a PV system as an imperative program using a fragment of decidable first-order theories~\cite{DBLP:books/daglib/0019162}.

@@

Our approach deals with the theoretical and pragmatic aspects of using model checking in stand-alone solar PV systems in order to tackle two issues: validation of a designed system and optimal sizing. Validation when a sized system must be evaluated in order to define if it meets the design requirements and user needs; and optimization when there are just the requirements and the PV system is not sized, or then the user tries to obtain a better solution that he or she has. We thus develop algorithms and the corresponding tools; moreover, we evaluate them using a comparative process \textcolor{red}{what do you mean by a ``comparative process''?} in several case studies with specialized simulation tools and real data from deployed PV systems.

@@

Note that it is out of our scope to perform code modification on the verifiers or the solvers used during the automated verification or automated synthesis steps of our approach. \textcolor{red}{why is it out of scope? You need to explain here.}

@@

\textcolor{red}{The beginning of this sentence is weird...}
Associated with \textbf{automated verification}, this thesis makes three main contributions. 
\textbf{First}, we propose an algorithm written in the C programming language, which implements the automated verification method to formally check the sizing and the operation of a given stand-alone PV system. 
\textbf{Second}, we evaluate the verification method by comparing three state-of-the-art model checkers in five real case studies. 
\textbf{Third}, experimental results show that this proposed approach can find subtle design errors in stand-alone PV systems, which are not easily detected by other approaches based on simulation. 
%
%The framework provided at this paper outlines the requirements, the process and the mathematical modeling to perform automated formal verification of a stand-alone PV system, using the satisfiability modulo theories (SMT)-based verification method.

\textcolor{red}{The second and third contributions are long and tough to read. I don't understand them.}
And linked to \textbf{automated synthesis}, our work makes more three original contributions: \textbf{First}, it is the first application of a sound and automated formal synthesis approach, which can provide accurate results of optimal sizing of stand-alone PV systems; 
\textbf{Secondly}, we propose a variant CEGIS method with striking differences of how the {\sc Synthesize} and {\sc Verify} phases work together, with the abolition of the feasible solution candidate vector and the use of an incremental, iterative loop to reach the optimal cost solution of the system; and 
\textbf{Thirdly}, experimental results with seven case studies show that formal synthesis approach qualitatively outperforms an existing state-of-the-art simulation tool. Our solution is far detailed and closer to the commercial reality (real PV systems) than the solution presented by simulation.

@@

The conversion of the traditional power grid into a smart grid, a fundamental example of a CPS, raises many issues that require novel methods and applications. In $2012$, a Chinese smart grid implementation was considered as case study to address the verification problem for performance and energy consumption~\cite{Yukseletall2012}. The authors employed a stochastic model checking approach and presented a modeling and analysis study using PRISM, which is a probabilistic model checker~\cite{KwiatkowskaNP11}. \textcolor{red}{This sentence is too long...} The focus of this study was on how CPSs integrate information and communication technology functions to the physical elements of a system for monitoring and controlling purposes; here, the authors employed automated verification of certain quantitative properties of the system, as probability of node failure in the long run, impact of repair service on the failure risk, and expected energy consumption, without any interest in power generation or even solar PV systems.

@@

\textcolor{red}{This sentence is unclear...}
~\autoref{chap:publications} presents a list of papers that were written during the thesis elaboration and the status of every submission until the date of this document.

>>>>

%\textcolor{red}{What is the relation between CPS and PV? We should make it clear already in the beginning...}
\textcolor{red}{Energy production, distribution, and optimization are all Cyber-Physical Systems (CPS) problems~\cite{UC}}. CPS are engineered systems, which are built from, and depend upon, the seamless integration of computational and physical components~\cite{NSF2015,ChavesIBCF19,iet-cps.2018.5006}. During operation, such components must frequently adapt to the operating environment changes (dynamics of the physical processes) faced at run-time. They must be able to continue to behave in a controlled and safe way, thus posing novel technical challenges for the software engineering of services and applications for CPS~\cite{Metzger2014}. Software pervasiveness in CPS places new challenges; in particular, highly dynamic environments, rapidly changing requirements, unpredictable and uncertain operating conditions demand new paradigms for software design~\cite{Filieri2015}.

While some research efforts do exist to enhance and optimize the software development processes for CPS, further investigation and discussion of better and more effective models are still needed in practice~\cite{Al-Jaroodi2016}. Among the opportunities for enhancements in the software development processes for CPS, there exists the need for developing new techniques and tools to support CPS requirements gathering and analysis to synthesize \textit{correct-by-construction} implementations of CPS. These techniques need to deal with predefined requirements enforced by nature and inherited constraints of the target CPS. Besides, they should be able to provide verification and validation mechanisms for completeness, correctness, and consistency~\cite{Al-Jaroodi2016}. Uncertainty and variability, at the same time, can be dealt with formal verification~\cite{NESSI}.

\textcolor{red}{Moving into the myriad of CPS systems, in this thesis we focus specifically on energy.} Lack of access to clean and affordable energy is considered a core dimension of poverty~\cite{Hussein2012}. Nevertheless, the world is progressing; in particular, the number of people without electricity access fell below 1 billion threshold for the first time in 2017~\cite{IEAweo2018}. The share of people without access to electricity from Africa is 58\%, while 19\% of the share comes from developing Asia, and 31\% from Latin America~\cite{IEAweo2018}. Numbers from Brazil show the aim to electrify 270 isolated places and 2.7 million people until 2023~\cite{EPE2018}. 
Furthermore, there exists a close relationship between the lack of energy and the low HDI (Human Development Index) of those localities~\cite{Coelho}. Therefore, increased access to energy allows economic growth and poverty alleviation \cite{Karekesi}.

@@

\textcolor{red}{The main objective of this thesis is thus to create a new approach of formal methods and evaluate it as an alternative to solve research issues from solar PV systems. We therefore foster the use of specific computer science techniques to tackle specific issues of PV systems, as design validation and sizing optimization.  In particular, we focus on automated verification and automated synthesis. Secondarily, we focus on: evaluating different state-of-art symbolic software verifiers (and solvers) with the goal of obtaining the best performance in our verification back-end; on applications written in ANSI-C that are platform-independent. %; and we do not perform code modification on the verifiers or solvers, we use it as available by their developers. 
We further evaluate our proposed solution by comparative with commercial simulation tool and by data collected from real PV systems deployed at riverside communities.}

%More specifically we will:
%The thesis produces a methodological research with innovative value regarding the first use of automated synthesis for optimal sizing of solar PV systems.

\textcolor{red}{This Ph.D. thesis has the novelty of create an automated verification approach to solar PV systems. The research led to the creation of two software to carry out computational experiments and to compare with commercial tools and real data from systems installed in the field. One software to perform automated verification and validate the solar PV systems behavior, and another one to produce the optimal sizing of solar PV systems through automated synthesis program.}

In this study, a mathematical model of each component of a stand-alone PV system, as panel solar, charge controller, batteries, inverter, and electrical load is used for formal verification. The behavior of each system component can be analyzed and observed with the support of those formal models, as a joint operation of the components, which in this case represents the operation of the solar PV system itself. The project requirements, as battery autonomy and power demand, besides weather conditions, as solar irradiance and ambient temperature, are provided to the proposed tool and automatically verified during the formal process. The model checking tool reports in which conditions a system does not meet the user requirements. A key benefit to this approach is that it helps in the detection of flaws in the design phase of system development, thereby considerably improving system reliability~\cite{Akram2018}. 
%Those programs are viewed as mathematical objects, and the use of 
%Bringing this technique to solar PV systems, i
%Instead of finding a bug or a software inconsistence, t
%Here it will be verified a stand-alone PV system, with all their components modeled mathematically,   
%Among the non-deterministic variables considerated, is it worth to mention environmental issues (like solar irradiance, and temperature), and the variation of the electrical load (power consumption) or the battery' state of charge, as shown in Fig.\ref{fig:flowchartgeneral}.

\textcolor{red}{The implementation of the proposed method is carried out by means of an algorithm implemented using the C programming language. We performed computing experimentation in three state-of-the-art model checkers to formally verify PV designs, aiming to evaluate performance and correctness: the C Bounded Model Checker (CBMC)~\cite{Kroening}, the Efficient SMT-based Bounded Model Checker (ESBMC)~\cite{esbmc2018}, and the Configurable Program Analysis Checker (CPAchecker).}
%three the efficient SMT-based bounded model checker (ESBMC)~\cite{esbmc2018}. %, which allows one to incrementally verify a PV system as an imperative program using a fragment of decidable first-order theories~\cite{DBLP:books/daglib/0019162}.

@@

Our approach deals with the theoretical and pragmatic aspects of using model checking in stand-alone solar PV systems in order to tackle two issues: validation of a designed system and optimal sizing. Validation when a sized system must be evaluated in order to define if it meets the design requirements and user needs; and optimization when there are just the requirements and the PV system is not sized, or then the user tries to obtain a better solution that he or she has. 
\textcolor{red}{We thus develop algorithms and the corresponding tools; moreover, we evaluate them with several case studies, in a comparative with specialized simulation tools, supported by real data from deployed PV systems.}

@@

Note that it is out of our scope to perform code modification on the verifiers or the solvers used during the automated verification or automated synthesis steps of our approach. \textcolor{red}{This thesis intend to create front-end applications to the verifiers and solvers (that it will be the back-end). That decision is based on the fact that we can be verifier-independent, and evaluate different back-ends, now and in the future, always searching for the best performance and soundness.}

@@

\textcolor{red}{Related with the proposed \textbf{automated verification} method, this thesis makes three main contributions.}
\textbf{First}, we propose an algorithm written in the C programming language, which implements the automated verification method to formally check the sizing and the operation of a given stand-alone PV system. 
\textbf{Second}, we evaluate the verification method by comparing three state-of-the-art model checkers in five real case studies. 
\textbf{Third}, experimental results show that this proposed approach can find subtle design errors in stand-alone PV systems, which are not easily detected by other approaches based on simulation. 
%
%The framework provided at this paper outlines the requirements, the process and the mathematical modeling to perform automated formal verification of a stand-alone PV system, using the satisfiability modulo theories (SMT)-based verification method.

And linked to \textbf{automated synthesis}, our work makes more three original contributions: \textbf{First}, it is the first application of a sound and automated formal synthesis approach, which can provide accurate results of optimal sizing of stand-alone PV systems; 
\textcolor{red}{\textbf{Secondly}, we propose a variant CEGIS method of synthesis with striking differences of how the {\sc Synthesize} and {\sc Verify} phases from original CEGIS work (without solution candidate vector and using an incremental and iterative loop to reach the optimal cost solution); and }
\textcolor{red}{\textbf{Thirdly}, experimental results with seven case studies show that our approach qualitatively outperforms state-of-the-art simulation tool: our solution is far detailed and closer to real PV systems than the solution presented by simulation.}

@@

The conversion of the traditional power grid into a smart grid, a fundamental example of a CPS, raises many issues that require novel methods and applications. In $2012$, a Chinese smart grid implementation was considered as case study to address the verification problem for performance and energy consumption~\cite{Yukseletall2012}. The authors employed a stochastic model checking approach and presented a modeling and analysis study using PRISM, which is a probabilistic model checker~\cite{KwiatkowskaNP11}. \textcolor{red}{The focus of this study was on how CPSs integrate information and communication technology functions to the physical elements of a system for monitoring and controlling purposes. The authors did not focus on power generation or even solar PV systems.}

@@

\textcolor{red}{~\autoref{chap:publications} presents a list of papers that we submitted to international journals and conference related to the use of automated verification method. It covers the years from 2015 to 2019, related to the Ph.D. process.}
